<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{rglobi vignette}
-->

rglobi vignette - an R interface to the aggregated biotic interaction data of GloBI (http://globalbioticinteractions.org)
======

### About the package

`rglobi` allows easy access to the GLoBI database by wrapping the existing API.  

********************

### Quick start

#### First, install `rglobi`


```{r}
install.packages("devtools")
devtools::install_github("rglobi", "ropensci")
```

```{r}
library("rglobi")
```

Note that since rglobi is still pretty new, only the dev version of the library is available at this time. We hope to publish a version to CRAN once the library matures. For more information see [GitHub](https://github.com/ropensci/rglobi).

#### Find interactions involving a certain species

Determining which species interact with each other (and how and where) is a major focus of ecology.  The Global Biotic Interactions Database offers data on multiple interaction types, and the `rglobi` library offers several ways to specify and access these interactions. `get_interactions()` is the primary function used if data on a specific interaction type is required.  A focus taxon is entered (may be specified as "Genus species" or higher level (e.g., Genus, Family, Class)).

```{r}
hsapiens <- get_interactions(taxon = "Homo sapiens", interaction.type = "preysOn")
head(hsapiens)
```

`get_predators_of()` and `get_prey_of()` are wrappers for `get_interactions()` that remove the need to specify interaction type

```{r}
hsapiens <- get_prey_of("Homo sapiens")
head(hsapiens)
```

For a complete list of supported interaction types, 

```{r}
get_interaction_types()
```

For data sources in which type of interactions was not specified, the interaction is labeled "interacts_with".

If you wish to view all interactions instead of specific types (e.g., parasitism and predation instead of just one of the two), the `get_interactions_by_taxa()` function allows this. In addition, the function provides options to search for interactions between two groups (source and target taxa, see above table) and only find interactions in a certain region.

```{r}
rattus <- get_interactions_by_taxa(sourcetaxon = "Rattus")
head(rattus)
```

Only a source taxa need be identified, but you can also specify a target taxon and/or geographic boundary (Coordinates must be in decimal degrees ([EPSG:4326](http://spatialreference.org/ref/epsg/wgs-84/)) and correspond to the west, south, east, and northern boundaries (i.e., min x, min y, max x, max y).  

```{r}
aves_crustacea_northern_hemisphere <- get_interactions_by_taxa( sourcetaxon = "Aves", targettaxon = "Crustacea", bbox=c(-180, 0, 180, 90 ))
head(aves_crustacea_northern_hemisphere)
```

#### Find interactions in a geographic areas

Instead of a taxon-specific approach, users may also wish to gather information on all interactions occuring in a specific area.  For example, a group developing ecoystem models for the Gulf of Mexico may want to consider all the data from that region.  `rglobi` enables this type of search by allowing users to specify a rectangular bounding box. Coordinates must be in decimal degrees ([EPSG:4326](http://spatialreference.org/ref/epsg/wgs-84/)) and correspond to the west, south, east, and northern boundaries (i.e., min x, min y, max x, max y).  

```{r}
gulfinteractions <- get_interactions_in_area( bbox=c(-97.0, 17.5, -81, 31))
head(gulfinteractions)
```

To see all locations for which interactions have entered in GloBi, 

```{r}
areas <- get_interaction_areas()
head(areas)
```
You can also restrict this search to a certain area:

```{r}
areas <- get_interaction_areas (bbox=c(-67.87,12.79,-57.08,23.32))
head(areas)
```

#### Custom Queries using Cypher

Currently, GloBI is powered by [neo4j](http://neo4j.org), a graph database. Most results created by the R functions provided by `rglobi` are basically wrappers for commonly used cypher queries. This way, you can still use the library without having to learn Cypher, a graph query language. However, if you feel adventurous, you can execute queries using Cypher only using the function `query()`:

```{r}
result <- query("START taxon = node:taxons(name='Homo sapiens') MATCH taxon<-[:CLASSIFIED_AS]-predator-[:ATE|PREYS_ON]->prey-[:CLASSIFIED_AS]->preyTaxon RETURN distinct(preyTaxon.name) as `prey.taxon.name` LIMIT 10")
result
```

This particular query returns the names of the first ten known prey or diet items of humans (_Homo sapiens_). More examples of GloBI specific Cypher queries can be found on the [GloBI Cypher Wiki](https://github.com/jhpoelen/eol-globi-data/wiki/cypher). 
